<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{An introduction to the spnetwork package}
\usepackage[utf8]{inputenc}
-->

# An introduction to the `spnetwork` package

The `spnetwork` R package provides a class and several methods for
handling spatial networks, where the edges (links) in the network
are represented by spatial lines and the nodes (vertices) by spatial
points. The spatial information (lines, points) is represented by the
`sp` package, the graph information (edges, vertices, direction)
is taken care of by the `igraph` package.  Spatial networks can
represent for instance road networks, or river networks.

The `spnetwork` allows the creation of spatial networks, plotting,
summarizing, subsetting, and conversion to lines, points, or graph
data structures. Also, weights of network edges (e.g. the length
of an edge, or the time it takes to pass it) can be set, retrieved,
and modified.

We will start with a little example that creates a toy network
from scratch. It looks a bit awkward; in practice we will read edge
data from external files, such as shapefiles. The following example
creates a `SpatialLines` object, with a single `Line` per feature:

```{r}
l0 = cbind(c(1, 2), c(0, 0))
l1 = cbind(c(0, 0, 0), c(0, 1, 2))
l2 = cbind(c(0, 0, 0), c(2, 3, 4))
l3 = cbind(c(0, 1, 2), c(2, 2, 3))
l4 = cbind(c(0, 1, 2), c(4, 4, 3))
l5 = cbind(c(2, 2), c(0, 3))
l6 = cbind(c(2, 3), c(3, 4))
library(sp)
L1 = function(l) Lines(list(Line(l)), as.character(substitute(l)))
sl = SpatialLines(list(L1(l0), L1(l1), L1(l2), L1(l3), L1(l4), L1(l5), L1(l6)))
```

The class `SpatialNetwork`, which looks like
```{r}
library(spnetwork)
showClass("SpatialNetwork")
```
derives directly from `SpatialLinesDataFrame`, and contains a slot
`g` of class `igraph`, and a slot `weightfield` that tracks which
information is used to specify weights. The remaining slots,
`lines`, `data`, `bbox` and `proj4string` are derived from
`SpatialLinesDataFrame`.

The function `SpatialNetwork` creates a spatial network from lines
only:
```{r}
sln = SpatialNetwork(sl)
summary(sln)
```
When given a set of lines only, as done here, it needs to sort out
which lines are connected by comparing all first and last points
of each line, and forms connections (shared nodes/vertices) when
they coincide.

Optionally, `SpatialNetwork` can be given an igraph object (in which
case it will not sort out the graph from the lines), information
to override the default weights (line lengths), and (as explained
below) also directions.

We can take out the graph constructed by `sln@g`, and plot several
of its properties:
```{r}
library(igraph)
g = sln@g
plot(V(g)$x, V(g)$y, col = V(g)$n, pch = 16, cex = 2, asp = 1)
lines(sl)
text(V(g)$x, V(g)$y, as.character(1:length(V(g))), pos = 4)
```
where node color indicates the number of edges it is shared by.

Here, `V(g)$x` is the `x` attribute of the graph vertices (`V`); similarly,
```{r}
E(g)$weight
```
retrieves graph edge attributes.

We can let `igraph` plot this graph, as in
```{r}
plot(g)
```

and see that x and y attributes are used for the layout, but axes
are rescaled and edges are replaced by straight lines.

# Directed graphs

In the example above, we can create a directed network by specifying
the link directions, with 0 indicating two-way, 1 indicating one-way
down-link (i.e. in the direction of the sequence of coordinates of a
line), and -1 indicating one-way up-link:
```{r}
sln = SpatialNetwork(sl, direction = c(0,0,1,1,-1,-1,-1))
summary(sln)
```
we now see that the number of graph edges (9) exceeds the number
of line segments (7), by a number equal to the number of two-way
streets (2). Each graph edge has an index indicating which lines (links)
correspond to it:
```{r}
g = sln@g
E(g)$link_index
```

The `igraph` plot of this graph now indicates edges direction:
```{r}
plot(g)
```

and also the `spnetwork` plot method can add arrows to links that
have only one direction:
```{r}
plot(sln, arrow_size = 1)
```

# A larger example: shortest path, network subset

We'll use Delauny triangulation from package deldir:
```{r}
library(deldir)
```

The following function converts a set of points into a
SpatialPolygons or into a SpatialLines object:

```{r}
dd <- function(x, ..., to = "polygons") {
    stopifnot(is(x, "Spatial"))
    cc = coordinates(x)
    dd = deldir(list(x = cc[, 1], y = cc[, 2]), ...)
    if (to == "polygons") {
        tm = triMat(dd)
        fn = function(i) Polygons(list(Polygon(rbind(cc[tm[c(i, i[1]),], ]))), i)
        SpatialPolygons(lapply(1:nrow(tm), fn), proj4string = x@proj4string)
    } else if (to == "lines") {
        segs = dd$delsgs
        fn = function(i) Lines(list(Line(cc[c(segs[i, 5], segs[i, 6]), ])), i)
        SpatialLines(lapply(1:nrow(segs), fn), proj4string = x@proj4string)
    } else stop("argument to should be polygons or lines")
}
```

We'll generate a set of 

```{r}
n = 300
```

points in a unit square:

```{r}
set.seed(5432)
pts = SpatialPoints(cbind(x = sort(runif(n)), y = runif(n)))
sl = dd(pts, to = "lines")
```

From this object, we can create a SpatialNetwork by

```{r}
net = SpatialNetwork(sl)
```

when plotted, we can again use colour to denote the number of vertices connected to each edge:

```{r}
plot(net)
points(net)
```

Now we compute the shortest path from the left-most point (1) to the right-most one (100):

```{r}
path = get.shortest.paths(net@g, 1, n, output = "both")
sp = as.vector(path$vpath[[1]])
ids = as_ids(path$epath[[1]])
```

and plot it

```{r}
plot(net, col = 'grey')
sel = net[ids,]
lines(sel, col = 'red', lwd = 2)
points(sel)
text(V(net@g)$x[c(1, n)], V(net@g)$y[c(1, n)], c("begin", "end"), pos = 4)
```

Note that `net[ids,]` subsets the `SpatialNetwork` object, including
the graph structure:  

```{r}
summary(sel)
```

As the edge weights are computed by Line
lengths, this is the geographically shortest path from start to end.

# Real networks: Toronto, Ottawa

Package `spnetwork` comes with two toy data sets: `torontocentre` and
`ottawacentre`.  In the following plot, the Toronto city centre is
plotted using green for two-way streets, and red for one-way streets,
and arrows indicating one-way road direction.

```{r}
data("torontocentre")
tc = SpatialNetwork(torontocentre, direction = torontocentre$DIRECTION_)
plot(tc, col = ifelse(torontocentre$DIRECTION_ == 0, 'green', 'red'), 
	arrow_size = .3, axes = TRUE)
```

# Shortest path through directed graphs

The following example first finds the shortest path between two
arbitrary points (orange) and the corresponding nearest points
formed by the graph nodes (blue). From these two blue points,
shortest paths are found and plotted by nodes going north (blue
circles) and going south (red triangles). Clearly, this leads to
two different routes in a directed network.

```{r}
data("ottawacentre")
oc = SpatialNetwork(ottawacentre, direction = ottawacentre$DIRECTION_)
par(mfrow = c(1, 2))
plot(oc, col = ifelse(ottawacentre$DIRECTION_ == 0, 'green', 'red'), arrow_size=.3)
plot(oc, col = ifelse(ottawacentre$DIRECTION_ == 0, 'green', 'red'))
# find two distant points:
m = rbind(
c( -75.70246, 45.40728),
c( -75.69669, 45.42286)
)
pts = SpatialPoints(m, CRS(proj4string(oc)))
points(pts, col = 'orange', pch = 16)
# find nearest nodes:
oc.p = as(oc, "SpatialPointsDataFrame")
nearest = apply(spDists(pts, oc.p), 1, which.min)
points(oc.p[nearest, ], col = 'blue', pch = 16)
path = get.shortest.paths(oc@g, nearest[1], nearest[2], output = "both")
sp = as.vector(path$vpath[[1]])
ids = as_ids(path$epath[[1]])
points(oc[,ids], col = 'blue', cex = .7, pch = 16)
path = get.shortest.paths(oc@g, nearest[2], nearest[1], output = "both")
# sp = as.vector(path$vpath[[1]])
ids = as_ids(path$epath[[1]])
points(oc[,ids], col = 'red', cex = 1, pch = 2)
```
