<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{SpatialNetwork demo}
\usepackage[utf8]{inputenc}
-->

# SpatialNetwork demo

```{r}
library(spnetwork)
library(sp)
library(igraph)

l0 = cbind(c(1, 2), c(0, 0))
l1 = cbind(c(0, 0, 0), c(0, 1, 2))
l2 = cbind(c(0, 0, 0), c(2, 3, 4))
l3 = cbind(c(0, 1, 2), c(2, 2, 3))
l4 = cbind(c(0, 1, 2), c(4, 4, 3))
l5 = cbind(c(2, 2), c(0, 3))
l6 = cbind(c(2, 3), c(3, 4))
LL = function(l, ID) Lines(list(Line(l)), ID)
l = list(LL(l0, "e"), LL(l1, "a"), LL(l2, "b"), LL(l3, "c"), LL(l4, "d"), LL(l5, "f"), LL(l6, "g"))
sl = SpatialLines(l)
sln = SpatialNetwork(sl)
```

```{r}
plot(sln@g$x, sln@g$y, col = sln@g$n, pch = 16, cex = 2, asp = 1)
lines(sl)
text(sln@g$x, sln@g$y, E(sln@g), pos = 4)
```

```{r}
plot(sln@g)
```

# A larger example: shortest path through a Delauny triangulations network

We'll use Delauny triangulation from package deldir:
```{r}
library(deldir)
```

The following function converts a set of points into a
SpatialPolygons or into a SpatialLines object:

```{r}
dd <- function(x, ..., to = "polygons") {
    stopifnot(is(x, "Spatial"))
    cc = coordinates(x)
    dd = deldir(list(x = cc[, 1], y = cc[, 2]), ...)
    if (to == "polygons") {
        tm = triMat(dd)
        fn = function(ix) {
            pts = tm[ix, ]
            pts = c(pts, pts[1])
            Polygons(list(Polygon(rbind(cc[pts, ]))), ix)
        }
        SpatialPolygons(lapply(1:nrow(tm), fn), proj4string = x@proj4string)
    } else if (to == "lines") {
        segs = dd$delsgs
        lst = vector("list", nrow(segs))
        for (i in 1:nrow(segs)) lst[i] = Lines(list(Line(cc[c(segs[i, 5], segs[i, 
            6]), ])), i)
        SpatialLines(lst, proj4string = x@proj4string)
    } else stop("argument to should be polygons or lines")
}
```

We'll generate a set of 100 points in a unit square:

```{r}
# set.seed(5432)
x = sort(runif(100)) # sort, so x[1] is left, x[100] is right
y = runif(100)
pts = SpatialPoints(cbind(x, y))
# Next, we'll get the SpatialLines object from it:

sl = dd(pts, to = "lines")
```
and plot it:

```{r}
plot(sl)
```

From this object, we can create a SpatialNetwork by

```{r}
ln = SpatialNetwork(sl)
```

when plotted, we can again use colour to denote the number of vertices connected to each edge:

```{r}
plot(ln@g$x, ln@g$y, col = ln@g$n, pch = 16, cex = 2, asp = 1)
lines(sl, col = "grey")
```

Now we compute the shortest path from the left-most point (1) to the right-most one (100):

```{r}
path = get.shortest.paths(ln@g, 1, 100, output = "both")
sp = as.vector(path$vpath[[1]])
ids = as_ids(path$epath[[1]])
```

and plot it

```{r}
plot(ln@g$x, ln@g$y, col = ln@g$n, pch = 16, cex = 1.5, asp = 1)
lines(sl, col = "grey")
points(ln@g$x[sp], ln@g$y[sp], col = "red", cex = 2)
text(ln@g$x[c(1, 100)], ln@g$y[c(1, 100)], c("start", "end"), pos = 4)
lines(ln[ids,], col = 'red', lwd = 2)
```

As the edge weights are computed by Line lengths, this is the geographically shortest path.
